---
title: "APLA Paper"
date: '`r Sys.Date()`'
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{APLA Paper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


## Upload Data and packages

```{r }


# Upload Packages for Map
library("tidyr")
library("dplyr")
library("ggplot2")
library("sf")
library("tmap")
library("rnaturalearth")
library("rnaturalearthdata")

# For color
library("RColorBrewer")
library("viridis")
library("ggrepel") # might be useful later

theme_set(theme_bw())
# Plot all titles in ggplot2 centered
theme_update(plot.title = element_text(hjust = 0.5))

library("APLA")

# Upload Data
#APLA_Data<- read.csv("Data_APLA.csv")         # Data APLA  
APLA_Data<- APLA::Data_APLA

# Upload Data for Refugee Numbers only
#APLA_Data_1<- read.csv("Data_APLA_1.csv")         # Data APLA  
APLA_Data_1<- APLA::Data_APLA_1  

```

## Refugees in Latin America

```{r}
# Check and change class variable refugees
class(APLA_Data_1$RefugeeAndLikeSit)
APLA_Data_1$RefugeeAndLikeSit<- as.numeric(APLA_Data_1$RefugeeAndLikeSit)

# Calculate maximum number of Refugees in each country (not used)
APLA_Data_1 %>%
group_by(Country) %>% summarize(m = max(RefugeeAndLikeSit))

# Create new variable All Other Countries
APLA_Data_Filtered_1<- APLA_Data_1 %>%
    filter(Country %in% c("Costa Rica","Ecuador","Guatemala","Honduras","Mexico","Venezuela","Rest of Region"))

# PLOT REFUGEE NUMBERS IN LATIN AMERICA
PLOT<- ggplot(APLA_Data_Filtered_1, aes(Year, RefugeeAndLikeSit, col= Country)) +
  geom_line()+
  facet_wrap(~ Country, scales = "free_y")+
  ggtitle("Refugees and People in Refugee-Like Situation in Latin America")+
  ylab("Refugee Numbers")+
  theme(plot.title = element_text("serif", size = "14", hjust = 0.5))+
  theme_bw()

# PLOT WITH ADJUSTED LABELS
require(scales)
PLOT + scale_y_continuous(labels = comma) + 
    scale_x_continuous(breaks = c(1990, 2000, 2010, 2018))+
    theme(legend.position = "none")
```

## Regulatory Complexity Over Time

```{r }
ggplot(APLA_Data, aes(Year, Regulatory_Complexity)) +
  geom_jitter(height=0.8, width = 0.7, aes(color="Country-Year"), alpha=0.5, size=2)+
  scale_colour_manual(name='', values = c("Locally Weighted Regression"= "red","Country-Year"="grey", "Linear Regression Line"= "blue"))+
  geom_smooth(linetype="dashed", aes(color="Locally Weighted Regression"), se=FALSE)+
  geom_smooth(method = 'lm', formula = y~x, aes(color="Linear Regression Line"), se=FALSE)+
  scale_shape(solid=FALSE)+
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2018))+
  theme_bw()+
  ggtitle("Regulatory Complexity of Asylum Policies in Latin America")+
  ylab("Regulatory Complexity")+
  theme(plot.title = element_text("serif", size = "14", hjust = 0.5))+
  theme(legend.title = element_blank())+
  theme(legend.key = element_rect(fill = "white")) + guides(color=guide_legend(override.aes=list(fill=NA))) #+
  #scale_color_brewer(palette="Dark2")

```

## Liberalisation Over Time

```{r }
ggplot(APLA_Data, aes(Year, Liberalisation)) +
  geom_jitter(height=0.3, width = 0.7, aes(color="Country-Year"), alpha=0.5, size=2)+
  scale_colour_manual(name='', values = c("Locally Weighted Regression"= "red","Country-Year"="grey", "Linear Regression Line"= "blue"))+
  geom_smooth(linetype="dashed", aes(color="Locally Weighted Regression"), se=FALSE)+
  geom_smooth(method = 'lm', formula = y~x, aes(color="Linear Regression Line"), se=FALSE)+
  scale_shape(solid=FALSE)+
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2018))+
  theme_bw()+
  ggtitle("Liberalisation of Asylum Policies in Latin America")+
  ylab("Liberalisation Score")+
  theme(plot.title = element_text("serif", size = "14", hjust = 0.5))+
  theme(legend.title = element_blank())+
  theme(legend.key = element_rect(fill = "white")) + guides(color=guide_legend(override.aes=list(fill=NA))) #+
  #scale_color_brewer(palette="Dark2")
```

## Maps of Development of Policy Over-Time Across Countries for Paper 1. Regulatory Complexity

```{r }
ggplot(APLA_Data, aes(Year, Regulatory_Complexity, col= Country)) +
  geom_line()+
  facet_wrap(.~ Country, ncol= 5)+
  theme(legend.position = "none")+  # to remove legend
  ggtitle("Regulatory Complexity in Asylum Policies across Latin America, 1990-2018") +
  ylab("Regulatory Complexity")
 
```

## Maps of Development of Policy Over-Time Across Countries for Paper 1. Liberalisation

```{r }
ggplot(APLA_Data, aes(Year, Liberalisation, col= Country)) +
  geom_line()+
  facet_wrap(.~ Country, ncol= 5)+
  theme(legend.position = "none")+  # to remove legend
  ggtitle("Liberalisation in Asylum Policies across Latin America, 1990-2018") +
  ylab("Liberalisation Score")
 
```

## Plotting Residuals

```{r }

Model1<-lm(formula=Regulatory_Complexity~Year, data=APLA_Data)

# Second step obtain presdicted and residual values
Model1$predicted <- predict(Model1)
Model1$residuals <- residuals(Model1)

# Eight Step  Change legend
ggplot(Model1, aes(x=Year, y=Regulatory_Complexity))+
  geom_jitter(width = 0.8, height = 0.2, color="lightgrey")+
  geom_smooth(method = "lm", se=FALSE, color= "lightgrey")+
  geom_segment(aes(xend=Year, yend=Model1$predicted), alpha= .2, color="lightgrey")+
  geom_point(aes(color=abs(Model1$residuals), size=abs(Model1$residuals)))+
  scale_color_continuous(low="black", high="red")+
  guides(color=FALSE, size=guide_legend("Size of Residuals"))+
  geom_point(aes(y=Model1$predicted), shape=1, color="lightgrey")+
  theme_bw()+
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2018))+
  ggtitle("Residuals of Regulatory Complexity in Asylum Policies in Latin America")+
  ylab("Regulatory Complexity")+
  theme(plot.title = element_text("serif", size = "14", hjust = 0.5))
```

## Regulatory Complexity vs Liberalisation

```{r }

APLA_Data_Filtered<- APLA_Data %>%
    filter(Regulatory_Complexity > 15)

ggplot(APLA_Data_Filtered, aes(Regulatory_Complexity, Liberalisation))+
  geom_jitter(width=0.5, height= 0.5)+
  geom_smooth()
```

## Map of Countries Codified APLA

```{r , fig.cap='Source: APLA Database'}
world <- ne_countries(scale = "medium", returnclass = "sf")   # Map of World
#class(world)

#APLA_Map <- read.csv("Data_APLA_Maps.csv")                     # Data APLA for Mapping 
APLA_Map <- APLA::APLA_Map

APLA_Sel<- APLA_Map %>%                                       # Create Database with Selected Years for Plotting
  filter(Year %in% c("1990", "2000", "2010", "2018"))

# Rename column where names is "name", so that I can merge with my other dataset
colnames(world)[4] <- "Country"

# VERY IMPORTANT TO MERGE MAP AND DATA 
Map_APLA_Data<- merge(APLA_Sel, world, by="Country")

# to transform MAP_APLA from data frame into sf and data.frame
st_geometry(Map_APLA_Data) <- Map_APLA_Data$geometry


# Display all countries codified so far + Compass + Scale Bar + labels changed!+ Title. USE THIS
tm_shape(Map_APLA_Data) + tm_borders("black", lwd= .5) + 
  tm_layout(title="1990-2018", title.size = 1.5, title.position =c(0.6, "top")) + 
  tm_polygons("Codified", title="Codified APLA Countries", palette= "Blues", style="fixed", breaks=c(0, 0.1, 1), 
              labels=c("Not Codified","Codified Countries")) + tm_compass(position = c(0.3, 0.35)) + 
  tm_scale_bar(width = 0.22, position = c(0.65, 0.08)) 
```
