---
title: "APLA Database"
author: "Omar Hammoud Gallego"
date: "19/11/2020"
output: ioslides_presentation

---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```



## Upload Packages and Data for Presentation 
```{r, Upload Packages}

# Upload Packages for Data analysis and Map
library("dplyr")
library("ggplot2")
theme_set(theme_bw())
library("sf")
library("tmap")
library("rnaturalearth")
library("rnaturalearthdata")
library("tidyr")
library("stringr")  # To do text analysis 

# For color
library("RColorBrewer")
library("viridis")
library("ggrepel") # might be useful later

# Plot all titles in ggplot2 centered
theme_update(plot.title = element_text(hjust = 0.5))

```

```{r, Upload Data}
# Set Working Directory
#setwd("C:/Users/omarh/Desktop/")

# Upload APLA Database
APLA<- APLA::APLA_Database 

# Upload Data for Refugee Numbers only
APLA_Data_1<- APLA::Data_Refugees_UNHCR         # Data APLA  

# Upload Data to Map All Countries Latin America
APLA_Map<- APLA::APLA_Map                      # Data APLA for Mapping 

```


```{r, Calculate Regulatory Complexity}

#APLA<- read.csv("APLA_Database.csv")
#summary(APLA)

# Transform as Tibble
APLA_T <- as_tibble(APLA)
#APLA_T

# Filter out columns on international agreements
APLA_T1<- APLA_T %>% select(1,2, 35:262)

# Only Questions of Liberalisation and Comments included
APLA_T2<- APLA_T1 %>% select(1,2, ends_with(c("2_1","3_1_1")))

# Select only commentary section to calculate number "Art" to calculate Regulatory Complexity
APLA_T3<- APLA_T2 %>%
  select(1,2, ends_with("3_1_1"))

# Delete of first row
APLA_T3= APLA_T3[-1,]

# Pivot Table for data manipulation
APLA_T_L <- pivot_longer(APLA_T3, cols = colnames(APLA_T3)[3:length(colnames(APLA_T3))],
                        names_to = "Question", values_to = "Comment")

# Transform variable from factor to string
APLA_T_L$Comment<- as.character(APLA_T_L$Comment)

# Value 1 if "Art" present "0" otherwise
APLA_T_L1<- APLA_T_L %>%
  mutate(Included= ifelse(grepl("Art", APLA_T_L$Comment), 1,0))

# Aggregate values of Included per Country, Year, and Question
APLA_T_L2<- APLA_T_L1 %>%
  group_by(Q1,Q2) %>%
  summarise(Frequency= sum(Included))

# Calculate Percentage by looking at number of "Included" per Country/Year as a % of total number of indicators = 57
APLA_T_L3<- APLA_T_L2 %>%
    mutate(Total=57) %>%
    mutate(Regulatory_Complexity= Frequency/Total*100) 

# Round Regulatory Complexity
APLA_T_L3$Regulatory_Complexity<- round(APLA_T_L3$Regulatory_Complexity)

# Rename Country and Year, and Year transformed in numeric
colnames(APLA_T_L3)[1] <- "Country"
colnames(APLA_T_L3)[2] <- "Year"
APLA_T_L3$Year<- as.numeric(as.character(APLA_T_L3$Year))
```

```{r, Calculate Liberalisation Variable}

# Run Regulatory Complexity Code First

# Only Questions of Liberalisation and Comments included
APLA_T4<- APLA_T1 %>% select(1,2, ends_with(c("2_1")))

# Delete of first row
APLA_T4= APLA_T4[-1,]

# Pivot Table for data manipulation
APLA_T_L4 <- pivot_longer(APLA_T4, cols = colnames(APLA_T4)[3:length(colnames(APLA_T4))],
                        names_to = "Question", values_to = "Liberalisation_Score")

# Mutate Liberalisation Score into numeric
APLA_T_L4$Liberalisation_Score<- as.numeric(as.character(APLA_T_L4$Liberalisation_Score, na.rm= TRUE))
APLA_T_L4$Question<- as.factor(APLA_T_L4$Question)

APLA_T_L4[complete.cases(APLA_T_L4),]

# Aggregate values of Included per Country, Year, and Question
APLA_T_L5<- APLA_T_L4 %>%
  na.omit() %>%
  group_by(Q1,Q2) %>%
  summarise(Liberal_Policies= sum(Liberalisation_Score == 0), Restrictive_Policies= sum(Liberalisation_Score == 1)) %>%
  mutate(Sum_Both = Liberal_Policies + Restrictive_Policies) %>%
  mutate(Liberalisation= Liberal_Policies/ Sum_Both)
  
# Round Liberalisation
APLA_T_L5$Liberalisation<- round(APLA_T_L5$Liberalisation, digits = 2)

# Take only values where Total_Both => 9
APLA_T_L6<- APLA_T_L5[APLA_T_L5$Sum_Both >= 9, ]

# Rename Country and Year, and Year transformed in numeric
colnames(APLA_T_L6)[1] <- "Country"
colnames(APLA_T_L6)[2] <- "Year"
APLA_T_L6$Year<- as.numeric(as.character(APLA_T_L6$Year))


```

## Plot Number of Refugees in Latin America, Regulatory Complexity and Liberalisation over Time for Presentation
```{r, Refugees in LA}
# Check and change class variable refugees
#class(APLA_Data_1$RefugeeAndLikeSit)
APLA_Data_1$RefugeeAndLikeSit<- as.numeric(APLA_Data_1$RefugeeAndLikeSit)

# Calculate maximum number of Refugees in each country (not used)
APLA_Data_1 %>%
group_by(Country) %>% summarize(m = max(RefugeeAndLikeSit))

# Create new variable All Other Countries
APLA_Data_Filtered_1<- APLA_Data_1 %>%
    filter(Country %in% c("Costa Rica","Ecuador","Guatemala","Honduras","Mexico","Venezuela","Rest of Region"))

# PLOT REFUGEE NUMBERS IN LATIN AMERICA
PLOT<- ggplot(APLA_Data_Filtered_1, aes(Year, RefugeeAndLikeSit, col= Country)) +
  geom_line()+
  facet_wrap(~ Country, scales = "free_y")+
  ggtitle("Refugees and People in Refugee-Like Situation in Latin America")+
  ylab("Refugee Numbers")+
  theme(plot.title = element_text("serif", size = "14", hjust = 0.5))+
  theme_bw()


```
## Plot Refugee Numbers
```{r, Plot Refugee Numbers}
# PLOT WITH ADJUSTED LABELS
require(scales)
PLOT + scale_y_continuous(labels = comma) + 
    scale_x_continuous(breaks = c(1990, 2000, 2010, 2018))+
    theme(legend.position = "none")

```

## Regulatory Complexity Over Time
```{r Regulatory Complexity Over Time}
ggplot(APLA_T_L3, aes(Year, Regulatory_Complexity)) +
  geom_jitter(height=0.8, width = 0.7, aes(color="Country-Year"), alpha=0.5, size=2)+
  scale_colour_manual(name='', values = c("Locally Weighted Regression"= "red","Country-Year"="grey", "Linear Regression Line"= "blue"))+
  geom_smooth(linetype="dashed", aes(color="Locally Weighted Regression"), se=FALSE)+
  geom_smooth(method = 'lm', formula = y~x, aes(color="Linear Regression Line"), se=FALSE)+
  scale_shape(solid=FALSE)+
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2018))+
  theme_bw()+
  ggtitle("Regulatory Complexity of Asylum Policies in Latin America")+
  ylab("Regulatory Complexity")+
  theme(plot.title = element_text("serif", size = "14", hjust = 0.5))+
  theme(legend.title = element_blank())+
  theme(legend.key = element_rect(fill = "white")) + guides(color=guide_legend(override.aes=list(fill=NA))) #+
  #scale_color_brewer(palette="Dark2")

```

```{r Liberalisation Over Time, eval = FALSE, include = FALSE}
ggplot(APLA_T_L6, aes(Year, Liberalisation)) +
  geom_jitter(height=0.3, width = 0.7, aes(color="Country-Year"), alpha=0.5, size=2)+
  scale_colour_manual(name='', values = c("Locally Weighted Regression"= "red","Country-Year"="grey", "Linear Regression Line"= "blue"))+
  geom_smooth(linetype="dashed", aes(color="Locally Weighted Regression"), se=FALSE)+
  geom_smooth(method = 'lm', formula = y~x, aes(color="Linear Regression Line"), se=FALSE)+
  scale_shape(solid=FALSE)+
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2018))+
  theme_bw()+
  ggtitle("Liberalisation of Asylum Policies in Latin America")+
  ylab("Liberalisation Score")+
  theme(plot.title = element_text("serif", size = "14", hjust = 0.5))+
  theme(legend.title = element_blank())+
  theme(legend.key = element_rect(fill = "white")) + guides(color=guide_legend(override.aes=list(fill=NA))) #+
  #scale_color_brewer(palette="Dark2")
```
## Regulatory Complexity Over Time Across Countries
```{r Plots of Development of Policy Over-Time Across Countries for Paper 1. Regulatory Complexity}
ggplot(APLA_T_L3, aes(Year, Regulatory_Complexity, col= Country)) +
  geom_line()+
  facet_wrap(.~ Country, ncol= 5)+
  theme(legend.position = "none")+  # to remove legend
  ggtitle("Regulatory Complexity in Asylum Policies across Latin America, 1990-2018") +
  ylab("Regulatory Complexity")
 
```

```{r Plots of Development of Policy Over-Time Across Countries for Paper 1. Liberalisation, eval = FALSE, include = FALSE}
ggplot(APLA_T_L6, aes(Year, Liberalisation, col= Country)) +
  geom_line()+
  facet_wrap(.~ Country, ncol= 5)+
  theme(legend.position = "none")+  # to remove legend
  ggtitle("Liberalisation in Asylum Policies across Latin America, 1990-2018") +
  ylab("Liberalisation Score")
 
```


## Plot Map of Countries Codified using APLA Dataset

```{r Map of Countries Codified APLA, fig.cap='Source: APLA Database'}
world <- ne_countries(scale = "medium", returnclass = "sf")   # Map of World
#class(world)

APLA_Sel<- APLA_Map %>%                                       # Create Database with Selected Years for Plotting
  filter(Year %in% c("1990", "2000", "2010", "2018"))

# Rename column where names is "name", so that I can merge with my other dataset
colnames(world)[4] <- "Country"

# VERY IMPORTANT TO MERGE MAP AND DATA 
Map_APLA_Data<- merge(APLA_Sel, world, by="Country")

# to transform MAP_APLA from data frame into sf and data.frame
st_geometry(Map_APLA_Data) <- Map_APLA_Data$geometry


# Display all countries codified so far + Compass + Scale Bar + labels changed!+ Title. USE THIS
tm_shape(Map_APLA_Data) + tm_borders("black", lwd= .5) + 
  tm_layout(title="1990-2018", title.size = 1.5, title.position =c(0.6, "top")) + 
  tm_polygons("Codified", title="Codified APLA Countries", palette= "Blues", style="fixed", breaks=c(0, 0.1, 1), 
              labels=c("Not Codified","Codified Countries")) + tm_compass(position = c(0.3, 0.35)) + 
  tm_scale_bar(width = 0.22, position = c(0.65, 0.08)) 
```

## Maps of Policy Measures Included in Legislation 1990 - 2018

```{r, Prepare data for plotting policy measures}
world <- ne_countries(scale = "medium", returnclass = "sf")
#class(world)

# Rename column where names is "name", so that I can merge with my other dataset
colnames(APLA)[1] <- "Name"
colnames(APLA)[2] <- "Year"
colnames(world)[4] <- "Name"
  
# VERY IMPORTANT TO MERGE MAP AND DATA 
Policy_Maps<- merge(APLA, world, by="Name")

# to transform IMPALA from data frame into sf and data.frame
st_geometry(Policy_Maps) <- Policy_Maps$geometry

# to add colour to map
library(RColorBrewer)
#display.brewer.all()

# Create Database with Selected Years for Plotting
Policy_Maps_Filtered<- Policy_Maps %>%                                       
  filter(Year %in% c("1990", "2000", "2010", "2018"))



```

## Map Refugee Definition
```{r, Maps Refugee Definition}
# Cartagena Refugee Definition LA7
tm_shape(Policy_Maps_Filtered) + tm_borders("black", lwd= .5) + tm_polygons("LA7.1_1", title="Cartagena Refugee Definition", palette= "Blues", style="fixed", breaks=c(0.0, 0.9, 1.0), labels=c("Not Incorporated","Cartagena Incorporated")) + tm_facets(by = "Year")

```

## Map Asylum in Constitution
```{r, Maps Asylum into Constitution}
# Asylum into Constitution  LA3
tm_shape(Policy_Maps_Filtered) + tm_borders("black", lwd= .5) + tm_polygons("LA3.1_1", title="Asylum into Constitution", palette= "Blues", style="fixed", breaks=c(0.0, 0.9, 1.0), labels=c("Not Included","Included into Constitution")) + tm_facets(by = "Year")
```

## Map First Country Asylum 
```{r, Map Asylum Constitution}

# First Country of Asylum  Q101
tm_shape(Policy_Maps_Filtered) + tm_borders("black", lwd= .5) + tm_polygons("Q101.1_1", title="First Country of Asylum Principle", palette= "Blues", style="fixed", breaks=c(0.0, 0.9, 1.0), labels=c("Not Included","Included into Legislation")) + tm_facets(by = "Year")
```

## Map Right to Work
```{r, Map Right to Work}
# Right to Work q54
tm_shape(Policy_Maps_Filtered) + tm_borders("black", lwd= .5) + tm_polygons("Q54.1_1", title="Right to Work", palette= "Blues", style="fixed", breaks=c(0.0, 0.9, 1.0), labels=c("Not Included","Included")) + tm_facets(by = "Year")
```

## Map No Penalisation for Illegal Entry 
```{r, Map No Penalisation for Illegal Entry }

# No Penalisation for Illegal Entry LA13
tm_shape(Policy_Maps_Filtered) + tm_borders("black", lwd= .5) + tm_polygons("LA13.1_1", title="No Penalisation for Illegal Entry", palette= "Blues", style="fixed", breaks=c(0.0, 0.9, 1.0), labels=c("Not Included","Included in Legislation")) + tm_facets(by = "Year")
```

## Map UNHCR represented in refugee status decision council
```{r, Map UNHCR represented in refugee status decision council}
# Is UNHCR represented in refugee status decision council LA53
tm_shape(Policy_Maps_Filtered) + tm_borders("black", lwd= .5) + tm_polygons("LA53.1_1", title="UNHCR representation in National Refugee Council", palette= "Blues", style="fixed", breaks=c(0.0, 0.9, 1.0), labels=c("Not Included","Included in Legislation")) + tm_facets(by = "Year")
```

## Map  Is UNHCR informed if negative last instance decision
```{r, Map  Is UNHCR informed if negative last instance decision}

# Is UNHCR informed if negative last instance decision
tm_shape(Policy_Maps_Filtered) + tm_borders("black", lwd= .5) + tm_polygons("LA57.1_1", title="UNHCR informed if Negative Decision", palette= "Blues", style="fixed", breaks=c(0.0, 0.9, 1.0), labels=c("Not Included","Included in Legislation")) + tm_facets(by = "Year")
```

## Map Reasonable Time Limit to Submit Appeal Request
```{r, Map Reasonable Time Limit to Submit Appeal Request}

# Reasonable Time Limit to Submit Appeal Request LA59
tm_shape(Policy_Maps_Filtered) + tm_borders("black", lwd= .5) + tm_polygons("LA59.1_1", title="Reasonable Time to Submit Appeal", palette= "Blues", style="fixed", breaks=c(0.0, 0.9, 1.0), labels=c("Less than 15 Days","15 Days or More")) + tm_facets(by = "Year")
```

## Map  Refugee Status due to Persecution based on Gender
```{r, Map  Refugee Status due to Persecution based on Gender}
# Refugee Status due to Persecution based on Gender LA65 
tm_shape(Policy_Maps_Filtered) + tm_borders("black", lwd= .5) + tm_polygons("LA65.1_1", title="Persecution due to Gender", palette= "Blues", style="fixed", breaks=c(0.0, 0.9, 1.0), labels=c("Not Included","Included in Legislation")) + tm_facets(by = "Year")


```
