---
title: "Report Americas UNHCR"
author: "Omar HG"
date: "06/01/2021"
bookdown::html_document2: default
---

```{r setup, include=TRUE, echo = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```


```{r, Upload Packages}

# Upload Packages for Data analysis and Map
library("dplyr")
library("ggplot2")
theme_set(theme_bw())
library("sf")
library("tmap")
library("rnaturalearth")
library("rnaturalearthdata")
library("tidyr")
library("stringr")  # To do text analysis 
library(rgeos)

library("bookdown")
# For color
library("RColorBrewer")
library("viridis")
library("ggrepel") # might be useful later

# Plot all titles in ggplot2 centered
theme_update(plot.title = element_text(hjust = 0.5))

```

```{r, Upload Data}
# Upload APLA Database
APLA<- APLA::APLA_Database 

# Upload Data for Refugee Numbers only
APLA_Data_1<- APLA::Data_Refugees_UNHCR         # Data APLA  

# Upload Data to Map All Countries Latin America
APLA_Map<- APLA::APLA_Map                      # Data APLA for Mapping 

```

```{r, Calculate Regulatory Complexity, echo = FALSE}

# Transform as Tibble
APLA_T <- as_tibble(APLA)
#APLA_T

# Filter out columns on international agreements
APLA_T1<- APLA_T %>% select(1,2, 35:262)

# Only Questions of Liberalisation and Comments included
APLA_T2<- APLA_T1 %>% select(1,2, ends_with(c("2_1","3_1_1")))

# Select only commentary section to calculate number "Art" to calculate Regulatory Complexity
APLA_T3<- APLA_T2 %>%
  select(1,2, ends_with("3_1_1"))

# Delete of first row
APLA_T3= APLA_T3[-1,]

# Pivot Table for data manipulation
APLA_T_L <- pivot_longer(APLA_T3, cols = colnames(APLA_T3)[3:length(colnames(APLA_T3))],
                        names_to = "Question", values_to = "Comment")

# Transform variable from factor to string
APLA_T_L$Comment<- as.character(APLA_T_L$Comment)

# Value 1 if "Art" present "0" otherwise
APLA_T_L1<- APLA_T_L %>%
  mutate(Included= ifelse(grepl("Art", APLA_T_L$Comment), 1,0))

# Aggregate values of Included per Country, Year, and Question
APLA_T_L2<- APLA_T_L1 %>%
  group_by(Q1,Q2) %>%
  summarise(Frequency= sum(Included))

# Calculate Percentage by looking at number of "Included" per Country/Year as a % of total number of indicators = 57
APLA_T_L3<- APLA_T_L2 %>%
    mutate(Total=57) %>%
    mutate(Regulatory_Complexity= Frequency/Total*100) 

# Round Regulatory Complexity
APLA_T_L3$Regulatory_Complexity<- round(APLA_T_L3$Regulatory_Complexity)

# Rename Country and Year, and Year transformed in numeric
colnames(APLA_T_L3)[1] <- "Country"
colnames(APLA_T_L3)[2] <- "Year"
APLA_T_L3$Year<- as.numeric(as.character(APLA_T_L3$Year))
```

```{r, Calculate Liberalisation Variable, echo = FALSE}

# Run Regulatory Complexity Code First

# Only Questions of Liberalisation and Comments included
APLA_T4<- APLA_T1 %>% select(1,2, ends_with(c("2_1")))

# Delete of first row
APLA_T4= APLA_T4[-1,]

# Pivot Table for data manipulation
APLA_T_L4 <- pivot_longer(APLA_T4, cols = colnames(APLA_T4)[3:length(colnames(APLA_T4))],
                        names_to = "Question", values_to = "Liberalisation_Score")

# Mutate Liberalisation Score into numeric
APLA_T_L4$Liberalisation_Score<- as.numeric(as.character(APLA_T_L4$Liberalisation_Score, na.rm= TRUE))
APLA_T_L4$Question<- as.factor(APLA_T_L4$Question)

APLA_T_L4[complete.cases(APLA_T_L4),]

# Aggregate values of Included per Country, Year, and Question
APLA_T_L5<- APLA_T_L4 %>%
  na.omit() %>%
  group_by(Q1,Q2) %>%
  summarise(Liberal_Policies= sum(Liberalisation_Score == 0), Restrictive_Policies= sum(Liberalisation_Score == 1)) %>%
  mutate(Sum_Both = Liberal_Policies + Restrictive_Policies) %>%
  mutate(Liberalisation= Liberal_Policies/ Sum_Both)
  
# Round Liberalisation
APLA_T_L5$Liberalisation<- round(APLA_T_L5$Liberalisation, digits = 2)

# Take only values where Total_Both => 9
APLA_T_L6<- APLA_T_L5[APLA_T_L5$Sum_Both >= 9, ]

# Rename Country and Year, and Year transformed in numeric
colnames(APLA_T_L6)[1] <- "Country"
colnames(APLA_T_L6)[2] <- "Year"
APLA_T_L6$Year<- as.numeric(as.character(APLA_T_L6$Year))


```

# Plot Map of Countries Codified using APLA Dataset

```{r Map of Countries Codified APLA, fig.cap='Source: APLA Database'}
world <- ne_countries(scale = "medium", returnclass = "sf")   # Map of World
#class(world)

APLA_Sel<- APLA_Map %>%                                       # Create Database with Selected Years for Plotting
  filter(Year %in% c("1990", "2000", "2010", "2020"))

# Rename column where names is "name", so that I can merge with my other dataset
colnames(world)[4] <- "Country"

# VERY IMPORTANT TO MERGE MAP AND DATA 
Map_APLA_Data<- merge(APLA_Sel, world, by="Country")

# to transform MAP_APLA from data frame into sf and data.frame
st_geometry(Map_APLA_Data) <- Map_APLA_Data$geometry


# Display all countries codified so far + Compass + Scale Bar + labels changed!+ Title. USE THIS
tm_shape(Map_APLA_Data) + tm_borders("black", lwd= .5) + 
  tm_layout(title="1990-2020", title.size = 1.5, title.position =c(0.6, "top")) + 
  tm_polygons("Codified", title="Codified APLA Countries", palette= "Blues", style="fixed", breaks=c(0, 0.1, 1), 
              labels=c("Not Codified","Codified Countries")) + tm_compass(position = c(0.3, 0.35)) + 
  tm_scale_bar(width = 0.22, position = c(0.65, 0.08)) 
```


To provide some context on the legal framework on asylum policies in Latin America, we present the results of a recent codification efforts of the asylum policies of 19 Latin American countries over the last 30 years (see Figure \@ref(fig: Map of Countries Codified APLA).  

As Figure \@ref(fig: Plots of Development of Policy Over-Time Across Countries for Paper 1. Regulatory Complexity) shows, over the last three decades countries across Latin America have reformed or adopted their asylum legislation. 

```{r Plots of Development of Policy Over-Time Across Countries for Paper 1. Regulatory Complexity}
ggplot(APLA_T_L3, aes(Year, Regulatory_Complexity, col= Country)) +
  geom_line()+
  facet_wrap(.~ Country, ncol= 5)+
  theme(legend.position = "none")+  # to remove legend
  ggtitle("Regulatory Complexity in Asylum Policies across Latin America, 1990-2020") +
  ylab("Regulatory Complexity")
 
```





# Detailed Policy Measures

```{r, Prepare data for plotting policy measures}
world <- ne_countries(scale = "medium", returnclass = "sf")
#class(world)

# Rename column where names is "name", so that I can merge with my other dataset
colnames(APLA)[1] <- "Name"
colnames(APLA)[2] <- "Year"
colnames(world)[4] <- "Name"
  
# VERY IMPORTANT TO MERGE MAP AND DATA 
Policy_Maps<- merge(APLA, world, by="Name")

# to transform IMPALA from data frame into sf and data.frame
st_geometry(Policy_Maps) <- Policy_Maps$geometry

# to add colour to map
library(RColorBrewer)
#display.brewer.all()

# Create Database with Selected Years for Plotting
Policy_Maps_Filtered<- Policy_Maps %>%                                       
  filter(Year %in% c("1990", "2000", "2010", "2020"))



```




While the scale and determinants of the on-going Venezuelan displacement crisis have been widely acknowledged to resemble the gravity of a refugee crisis, in few cases countries in the regions have adopted their legal frameworks for the reception of Venezuelan nationals.  
